---
title: "Practical Machine Learning"
output: html_notebook
---
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).



```{r include=FALSE, cache=FALSE}
library(caret)
library(randomForest)
library(rpart)
```


We first load in he data. The csv-files have been stored locally. 
```{r}
set.seed(1234)
datasetTrain <- read.csv("Data/pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
datasetTest <- read.csv('Data/pml-testing.csv', na.strings=c("NA","#DIV/0!", ""))

datasetTrain<-datasetTrain[,colSums(is.na(datasetTrain)) == 0]
datasetTest <-datasetTest[,colSums(is.na(datasetTest)) == 0]


datasetTrain   <-datasetTrain[,-c(1:7)]
datasetTest <-datasetTest[,-c(1:7)]


dim(datasetTrain)
dim(datasetTest)
print (paste("Test Set Length:        ", length(datasetTest)))
print (paste("Training Set Length:    ", length(datasetTrain)))


```


Here we split up the dataset into subsets. 75% for training and 25% for testing.
```{r}
splitPct = 0.75


subset <- createDataPartition(y=datasetTrain$classe, p=splitPct, list=FALSE)
subsetTraining <- datasetTrain[subsamples, ] 
subsetTesting <- datasetTrain[-subsamples, ]
dim(subsetTraining)
dim(subsetTesting)
print (paste("Test Set Length:        ", length(subsetTesting)))
print (paste("Training Set Length:    ", length(subsetTraining)))
```



Here we do the actual computation. Notice that we use doParallel which only works for som algorithms.
The computation has also been done with LSA and GBM but we have finally chosen Random Forest as the algorithm because of better results.
```{r}
useParallel = FALSE

method = "randomforest"

cpu_cores = detectCores(all.tests = FALSE, logical = TRUE)
print(paste("CPU Cores:" , cpu_cores))


if(useParallel) {
  print ("Using Parallel")
  library(doParallel)
  cl <- makeCluster(detectCores())
  registerDoParallel(cl)
}

if(method == "randomforest") {
  print("Computing Random Forest")
  model <- randomForest(classe ~. , data=subsetTraining, method="class")
} else if(method == "train_randomforest") {
  print("Computing Random Forest (train)")
  model <- train(classe ~ ., data=datasetTrain, method="rf")
} else {
  print("Unknown Mehtod!")
}



prediction <- predict(model, subsetTesting, type = "class")

confusionMatrix(prediction, subsetTesting$classe)

if(useParallel) {
  stopCluster(cl)
}
```

Accuracy for the random forest algorithm is 99.5%.
Below we show prediction for the 20 cases:
```{r}
predictfinal <- predict(model, datasetTest, type="class")
predictfinal
```

